High-Throughput Phenotyping using Computer Vision and Machine Learning \\ \large SMC Data Challenge \#1

### 2023-07-28 paper/main.tex
final draft

%%
%% Copyright 2007-2020 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

%\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
 \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{caption}
\usepackage{hyperref}

%% Hide "preprint submitted by" line
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

%% You might want to define your own abbreviated commands for common used terms, e.g.:
\newcommand{\kms}{km\,s$^{-1}$}
\newcommand{\msun}{$M_\odot}

@@

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={},
%%            city={},
%%            postcode={},
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{%
High-Throughput Phenotyping of Populus Trichocarpa Using Computer Vision \\
\large SMC Data Challenge \#1
}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[]{Vivaan Singhvi}
\author[]{Langalibelele Lunga}
\author[]{Pragya Nidhi}
\author[]{Chris Keum}
\author[]{Varrun Prakash}

\begin{abstract}
%% Text of abstract
	High-throughput phenotyping refers to the non-destructive and efficient evaluation of plant phenotypes. Machine Learning methods offer the prospect of addressing two major challenges in phenotyping plants: efficiency in handling large datasets and extraction of specific traits. Previous techniques, such as the application of Deep Neural Nets in tandem with automated cameras, to our best knowledge, refrain from using physically labeled plants to organize datasets. To accomplish this, we used Optical Character Recognition(OCR) to read the physical labels on the plants, image segmentation techniques(OpenCv, FacebookSAM)  in order to apply Machine Learning algorithms(RandomForestClassifier, XGradientBoosting) for plant morphology classifications, and analyzed encoded EXIF tag information for the purpose of finding correlations between phenotypes. We used a  Populus Trichocarpa  plant dataset provided by Oak Ridge National Laboratory, with 1,672 image files of plants labeled with treatment( Control or Drought), block, row, position, and genotype. We found that OCR techniques, such as PaddleOCR, successfully read 91.3\% of the labeled plants, allowing for labeled information to be accurately placed in a pandas Dataframe. The success of the OCR carried into our XGradientBoosting and RandomForestClassifier techniques, which classified leaf morpholgoies by leaf shape and color with an accuracy of 62.82\%. Finally, we found [insert findings from step 4]. The use of Machine Learning in High-throughoutPut Phenotyping has shown to be effective in analyzing large plant datasets[Improve this sentence]. It allows for efficient and environmentally friendly classification of plants.
\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword, up to a maximum of 6 keywords
computer vision \sep machine learning \sep high-throughput phenotyping

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

@@

Image-based phenotyping refers to a breakthrough technology used in plant biology and agriculture to examine and assess plants’ “anatomical, ontological, physiological, and biochemical features” through the use of images. Previous studies have shown its potential as a noninvasive replacement for traditional on field techniques used to extract important phenotypic data. In recent years, this potential has been further intensified by coupling image-based phenotyping techniques with machine learning algorithms. This new approach, called high throughput phenotyping, has enabled the extraction of phenotypes from “complex” plant image datasets that were previously challenging to analyze efficiently. Furthermore, more recent studies have shown that the technology holds potential in identifying correlations between phenotype, genotype, and environmental metadata.

In this study, we aim to develop an accurate Machine Learning Model for High Throughput Phenotyping of leaves’ morphological traits (e.g. leaf size, shape, color) in plant images; unlike datasets in previous research, the dataset used in this study contains physical labels within the images, each containing important information on treatment (Control or Drought), block, row, position, and genotype. As a result, this study will deal with a new aspect of image-based phenotyping namely, extracting data from the white labels to identify correlations between leaves’ phenotypes and data embedded within the white labels. Given this information, the study aims to answer the following questions:

@@

\item Is it possible to use optical character recognition (OCR) or machine learning techniques to “Read” the label on each tag and generate a spreadsheet containing the treatment, block, row, position, and genotype? Doing this would dramatically simplify data collection, as this information is usually collected manually.

@@

Traditionally, plant morphologies would have to be analyzed by hand, wasting time and potentially damaging the environment. Now, with breakthroughs in high throughput image phenotyping led by Machine Learning techniques, methods such as Automated Machine Learning \citep{koh2021automated}, Computer Vision with Deep Learning \citep{mochida2019computer},  and Convolutional Neural Networks(CNNs) \citep{pound2017deep}, have allowed for efficient and timely analysis of plant phenotypes. With these methods, researchers are able to extract features from plants and classify plant morphologies effectively with little effect on the plant environment.

@@

Advancements in the field of plant phenotyping have been pioneered by technology. 
From using hardware, such as Raspberry Pi imaging systems \citep{tausen2020greenotyper} to thermographical sensing techniques \citep{walter2015plant}, research in the field of high phenotyping has led to an acceleration in working efficiency. Plants are now able to be analyzed in fast, effective, and accurate methods allowing for plant to be used in scientific discoveries. In our study, we rely on machine learning techniques to classify morphologies based on a phone image dataset provided.

@@

Machine Learning has enabled researchers with computer vision, allowing them to create programs that “read” image information in real time. In addition, it increases the efficiency of obtaining leaf morphology data. However, researchers have found that the use of traditional classification techniques, such as RandomForrestClassifier, have a high performance, but do not generalize well across datasets \citep{pound2017deep}. In order to increase accuracy of models and the generalization of the model, Deep Learning techniques have been introduced. With deep learning models, models train themselves iteratively until they reach a desired outcome. These techniques, such as Convolutional Neural Networks \citep{koh2021automated, pound2017deep} have shown to be highly accurate in classifying plant structures while maintaining performance across different datasets. Convolutional Networks consist of several layers, which allow for more discriminative and detailed analysis of the plant image, leading to highly accurate classifications. In our study, we use traditional methods(GradientBoosting, RandomForest) to classify plant morphology, as our focus is on organizing the plant images into an easily readable dataset and performing classifications, instead of creating models highly transferable across datasets.

@@

Due to global events, such as climate change and global population increase, the ability to produce large amounts of healthy food will be crucial to society. With Machine Learning being used for phenotyping, researchers will be able to rapidly analyze food crops to maximize production and crop breeding \citep{arya2022deep, shakoor2017high}.
As climate change and global population increases, global crop yield will have to increase to provide for the growing populations. Further research has shown that climate change will continue to negatively affect crops and cause crop diseases \citep{newton2011implications}, hence the need for effective and precise analysis of plants, which is done with Machine Learning models. In our study, we aim to show how Machine Learning can be used to organize plant data, dissect plant images into meaningful classifications, and allow for rapid investigation of plants for scientists in the field of plant phenotyping.

@@

Given the nature of the project, there are many aspects requiring obtaining large files, such as the dataset itself or large models. Therefore, all scripts needed to fully set up the project (downloading large files or setting up a large virtual environment), are viewable in the \verb|scripts/| folder in the GitHub repository, linked in Section \ref{code}.

@@

@@

To read the text, a pre-trained Optical Character Recognition (OCR) model seemed like the optimal choice. However, there are several models high-performing models available for use. The three candidates chosen for this project were \verb|PyTesseract| \citep{lee2017pytesseract}, \verb|EasyOCR| \citep{jaided2020easyocr}, and \verb|PaddleOCR| \citep{du2020ppocr}.

@@

After the text was read by the OCR, regular expressions (RegEx) were used to find the precise details mentioned in the dataset: treatment, block, row, position, and genotype. A subjective analysis of the images in the dataset showed that the treatment was either a ‘C’ or a ‘D’, and the remaining features followed the RegEx patterns below (note: \verb|\d| represents a digit, anything between 0-9) :

@@

After all the text is read and processed, it is converted to a Pandas \verb|DataFrame| and saved to an Excel file.

@@

In order to be able to determine the morphological characteristics of the plants, it is necessary to perfectly segment the leaves from each image. This proved to be a challenging endeavor due to the complexity of the background of the images, other datasets, such as the Komatsuna dataset, \citep{uchiyama2017easy} had 'cleaner' image backgrounds.

To obtain perfect masks for leaves in each image, the Segment Anything Model (SAM) \citep{kirillov2023segment} is the obvious choice. Being a relatively new and groundbreaking model, this paper is one of the first to utilize a pre-trained model of its caliber.

@@

\begin{center}
	\centering
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/1.png}	
        \includegraphics[width=0.15\textwidth]{images/seg_pipeline/2.png}      	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/3.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/4.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/5.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
	\captionof{figure}{Pipeline of image transformations for leaf approximation, read left to right}
	% \label{fig_mom0}%
\end{center}

@@

Thus, it was necessary to train a machine learning model that was able to accurately detect these kinds of 'bad' leaves. Training data was generated by showing randomly selected leaves and getting human input regarding their suitability, extracting their features, and storing the data in a file. The model was implemented using Scikit-Learn's \verb|RandomForestClassifier|. The model had an accuracy of 90.91\% on testing data generated by Scikit-Learn's \verb|train_test_split|.

\begin{center}
	\centering
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
        \includegraphics[width=0.15\textwidth]{images/seg_filter/1.png}      	\includegraphics[width=0.15\textwidth]{images/seg_filter/2.png}	
	\captionof{figure}{The process for filtering segmentations through the model: from contour approximations to generated masks to filtered masks}
	% \label{fig_mom0}%
\end{center}

Then, each leaf could be cropped and rotated, which, when combined with masking every pixel outside the its border, effectively highlighted and isolated it for visual purposes. By repeating this process for each image in the dataset, random leaves could then be selected to generate data for training the morphological classification model.

By using a slightly modified version of the aforementioned feature extraction function in conjunction with human labeling of leaves regarding their shape, color, and level of brown splotches, a \verb|MultiOutputClassifier| was able to be implemented with Scikit-Learn. More specifically, the \verb|XGBoost| algorithm was used for each of the three predicted features in the \verb|MultiOutputClassifier|.

TO BE CONTINUED (REWRITING NOT FINISHED YET).

@@

The challenge for using classification data as features to train a model with is that they are categorical, rather than quantitative variables; thus, extra transformations must be done to properly represent the data.

To properly represent the classifications, the method of one-hot encoding (MAYBE FIND SOURCE SAYING ITS GOOD) was implemented using the Pandas \verb|get_dummies| method. This is a process in which a feature is split into multiple columns, with each row having a one if it matches the column or a zero if it doesn't.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{c||c c c c} 
 \hline
 leaf\_color & light\_green & dark\_green & yellow\_green & yellow  \\
 \hline
 light\_green & 1 & 0 & 0 & 0 \\
 yellow\_green & 0 & 0 & 1 & 0 \\
 dark\_green & 0 & 1 & 0 & 0 \\ 
 light\_green & 1 & 0 & 0 & 0 \\
 yellow & 0 & 0 & 0 & 1 \\
 \hline
\end{tabular}
}
\captionof{table}{Example transformation of a column with one-hot encoding, with the original column on the left and the encoded columns on the right}
\label{Table1}
\end{center}

This process was done with the leaf color and shape variables. However, since the level of brown splotches is ordinal, it was assigned levels 0, 1, 2, and 3, corresponding to ‘none’, ‘low’, ‘medium,’ and ‘high.’ Now, with all the data numerical, it could be used to train and test the model.

@@

Firstly, we aimed to measure the size of a leaf through metadata, encoded in EXIF Tags, which are embedded within each of the images. These EXIF tags typically give important information about the images, including the type of camera utilized, the distance of the camera from a photo, and photo dimensions (in pixels). In order to initiate the extraction of such metadata from the images, a package called “exif” was used and gave a variety of retrieval options.

Out of all the metadata extracted from the EXIF tags, the most notable pieces of information included ResolutionX/Y tags and focal length. Although these pieces of information are crucial in developing a mechanism to measure leaf size in the dataset, the information provided simply does not suffice in determining leaf size; we needed information such as how far the leaf was from the camera, information that was not given from EXIF tags alone. This lack of information rendered the metadata as inadequate for direct measurement of leaf size.

Another approach we considered to measure leaf size included using the white labels, held up by the researcher in each image, as a known reference object. Although this method seemed promising, there were certain limitations in most images in the dataset that prevented a consistent measurement. Firstly, the white label in the image was often obstructed by the researcher’s hands, as shown in Figure 1. This obstruction would ultimately interfere with alignment of the white label, making the label unreliable as an object of reference. Another limitation to this method came from the fact that the labels were often tilted, not in the same plane as the leaves. The tilt of the labels caused inaccuracies in measurement and would lead to misrepresentation of the true length. Along with that, there was no way of finding out how far forward or behind the leaf was relative to the tag, meaning the method would result in an inaccurate representation of the leaf.

In the second part of challenge 4, we aimed to find correlation among phenotypes using geographical features including soil type and weather conditions. Using the EXIF tags, we were able to retrieve the coordinates where the image was taken. We noticed all of the images contained were taken in approximately the same area, meaning weather would be the same for all plants. Along with that, most soil map APIs contain information about the soil in an area, so it would be unable to distinguish miniscule changes in locations. For all other potential geographical changes, due to the close proximity all images are taken in, we were unable to find out if there were any correlations. Not only was the location the same, the images were taken in a short time frame, within a week. This gave us the same limitations as the close location had.

\section{Results}
%%\label{}

@@

@@

>>>>

\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

\usepackage{caption}
\usepackage[hidelinks]{hyperref}

\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother

@@

\title{
High-Throughput Phenotyping using Computer Vision and Machine Learning \\
\large SMC Data Challenge \#1
}

\author[]{Vivaan Singhvi}
\author[]{Langalibalele Lunga}
\author[]{Pragya Nidhi}
\author[]{Chris Keum}
\author[]{Varrun Prakash}

\begin{abstract}
%% Text of abstract
High-throughput phenotyping refers to the non-destructive and efficient evaluation of plant phenotypes. In recent years, it has been coupled with machine learning in order to improve the process of phenotyping plants by increasing efficiency in handling large datasets and developing methods for the extraction of specific traits. Previous studies have developed methods to advance these challenges through the application of deep neural networks in tandem with automated cameras; however, the datasets being studied often excluded physical labels. In this study, we used a dataset provided by Oak Ridge National Laboratory with 1,672 images of Populus Trichocarpa with white labels displaying treatment (control or drought), block, row, position, and genotype. Optical character recognition (OCR) was used to read these labels on the plants, image segmentation techniques in conjunction with machine learning algorithms were used for morphological classifications, machine learning models were used to predict treatment based on those classifications, and analyzed encoded EXIF tags were used for the purpose of finding leaf size and correlations between phenotypes. We found that our OCR model successfully read 91.3\% of the labeled plants, allowing for the information to be accurately placed in a spreadsheet. Our classification models identified leaf shape, color, and level of brown splotches with an average accuracy of 62.82\%, and plant treatment with an accuracy of 60.08\%. Finally, we identified a few crucial pieces of information absent from the EXIF tags that prevented the assessment of the leaf size. There was also missing information that prevented the assessment of correlations between phenotypes and conditions. However, future studies could improve upon this to allow for the assessment of these features. The use of machine learning and computer vision in high-throughput phenotyping has shown to be effective in analyzing large plant datasets, leading to a more comprehensive phenotype analysis in plants and showing potential in various agricultural and environmental applications.
\end{abstract}

\begin{keyword}
computer vision \sep machine learning \sep high-throughput phenotyping
\end{keyword}

@@

\subsection{Background Information}
\label{background}

High-throughput phenotyping is defined as a breakthrough technology used in plant biology and agriculture to examine and assess plants' "anatomical, ontological, physiological, and biochemical features" through the use of images. Previous studies have shown its potential as a noninvasive replacement for traditional on-field techniques used to extract important phenotypic data. In recent years, this potential has been further intensified by coupling image-based phenotyping techniques with machine learning algorithms. This new approach has enabled the extraction of phenotypes from "complex" plant image datasets that were previously challenging to analyze efficiently. Furthermore, newer studies have shown that the technology has potential to identifying correlations between phenotype, genotype, and environmental metadata.

\subsection{Research Objective}
\label{objective}

In this study, we aim to develop an accurate machine learning model for high-throughput phenotyping of leaves’ morphological traits (e.g. leaf size, shape, color) in plant images; unlike datasets in previous research, the dataset used in this study contains physical labels within the images, each containing important information on treatment (control or drought), block, row, position, and genotype. As a result, this study will address a new aspect of image-based phenotyping, namely, extracting data from the white labels to identify correlations between leaves’ phenotypes and data embedded within the white labels. During this project, we aim to answer the following questions:

@@

\item Is it possible to use optical character recognition (OCR) or machine learning techniques to “read” the label on each tag and generate a spreadsheet containing the treatment, block, row, position, and genotype? Doing this would dramatically simplify data collection, as this information is usually collected manually.

@@

Traditionally, plant morphologies would be analyzed manually, wasting time and potentially damaging the environment. However, current breakthroughs in high-throughput phenotyping led by machine learning techniques\citep{koh2021automated, pound2017deep} and computer vision with deep learning \citep{mochida2019computer} have allowed for efficient and timely analyses of plant phenotypes. With these methods, researchers can extract features from plants and classify their morphologies effectively while having little effect on the plant itself.

@@

Advancements in the field of plant phenotyping have been pioneered by technology. From using hardware, such as Raspberry Pi imaging systems \citep{tausen2020greenotyper} to thermographic sensing techniques \citep{walter2015plant}, research in the field of high-throughput phenotyping has led to an acceleration in working efficiency. Plants are now able to be analyzed in fast, effective, and accurate methods allowing for them to be used in scientific discoveries. In our study, we rely on machine learning techniques to classify morphologies based on a phone image dataset provided.

@@

Machine learning has enabled researchers with computer vision, allowing them to create programs that “read” image information in real-time. In addition, it increases the efficiency of obtaining leaf morphology data. However, researchers have found that traditional classification techniques, such as Random Forest Classification, have high performance but do not generalize well across datasets \citep{pound2017deep}. In order to increase the accuracy of models and the generalization of the model, deep learning techniques have been introduced. With deep learning models, models train themselves iteratively until they reach a desired outcome. These techniques, such as convolutional neural networks \citep{koh2021automated, pound2017deep} have shown to be highly accurate in classifying plant structures while maintaining performance across different datasets. Convolutional neural networks consist of several layers, which allow for a more discriminative and detailed analysis of plant images, leading to highly accurate classifications. In our study, we use traditional methods to classify plant morphology, as our focus is performing classifications with a single dataset rather than creating models highly transferable across datasets.

@@

Due to global events, such as climate change and global population increase, the ability to produce large amounts of healthy crops will be crucial to society. With machine learning being used for phenotyping, researchers will be able to rapidly analyze food crops to maximize production and breeding \citep{arya2022deep, shakoor2017high}.

As climate change and the global population increase, global crop yield will have to increase to provide for the growing populations. Further research has shown that climate change will continue to negatively affect crops and cause crop diseases \citep{newton2011implications}, hence the need for effective and precise analysis of plants, which can be done with machine learning models. In our study, we aim to show how machine learning can be used to organize plant data, dissect plant images into meaningful classifications, and allow for rapid investigation of plants for scientists in the field of plant phenotyping.

@@

Given the nature of the project, there are many aspects requiring obtaining large files, such as the dataset itself or large models. Therefore, all scripts needed to fully set up the project (downloading large files or setting up a large virtual environment), are viewable in the \verb|scripts/| folder of the GitHub repository, linked in Section \ref{code}.

@@

All models used in this project are trained and tested on data split by Scikit-Learn's \verb|train_test_split|. Additionally, Scikit-Learn's \verb|accuracy_score| is used to determine the accuracy of said models.

@@

To read the text, a pre-trained Optical Character Recognition (OCR) model seemed like the optimal choice. However, there are several high-performing models available for use. The three candidates chosen for this project were \verb|PyTesseract| \citep{lee2017pytesseract}, \verb|EasyOCR| \citep{jaided2020easyocr}, and \verb|PaddleOCR| \citep{du2020ppocr}.

@@

After the text was read by the OCR, regular expressions (RegEx) were used to find the precise details mentioned in the dataset: treatment, block, row, position, and genotype. A subjective analysis of the images in the dataset showed that the treatment was either a \verb|C| or a \verb|D|, and the remaining features followed the RegEx patterns below (note: \verb|\d| represents a digit, anything between 0-9) :

@@

After all the text is read and processed, it is converted to a Pandas \verb|DataFrame| and saved to an Excel spreadsheet, which will be used in the following steps. Then, 30 random images are selected and the accuracy of the results is tested by comparing inputted to read features. As with the OCR testing earlier, it is calculated both with and without unread values.

@@

In order to be able to determine the morphological characteristics of the plants, it was necessary to perfectly segment the leaves from each image. This proved to be a challenging endeavor due to the complexity of the background of the images; other datasets, such as the Komatsuna dataset, \citep{uchiyama2017easy} had 'cleaner' image backgrounds.

To obtain perfect masks for leaves in each image, the Segment Anything Model (SAM) \citep{kirillov2023segment} was the obvious choice, due to its groundbreaking accuracy and ease of use. Being a relatively new model, this paper is one of the first to utilize a pre-trained model of its caliber.

@@

\begin{center}
	\centering
	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/1.png}	
        \includegraphics[width=0.23\textwidth]{images/seg_pipeline/2.png}      	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/3.png}	
	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/4.png}	
	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/5.png}	
	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/6.png}	
	\captionof{figure}{Pipeline of image transformations for leaf approximation, read left to right. The last image visualizes detected contours}
	% \label{fig_mom0}%
\end{center}

@@

Thus, it was necessary to train a machine learning model that was able to accurately detect these kinds of 'bad' leaves. Training data was generated by showing randomly selected leaves and getting human input regarding their suitability, extracting their features, and storing the data in a file. The model was implemented using Scikit-Learn's \verb|RandomForestClassifier|, and had an accuracy of 90.91\% on testing data.

\begin{center}
	\centering
	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/6.png}	
        \includegraphics[width=0.23\textwidth]{images/seg_filter/2.png}	
	\captionof{figure}{From leaf approximations with image processing to filtered leaf segmentations, from the SAM/ONNX Runtime (green leaves fit for classification)}
	% \label{fig_mom0}%
\end{center}

Then, each leaf could be cropped and rotated, which, when combined with masking every pixel outside its border, effectively highlighted and isolated it for visual purposes. By repeating this process for each image in the dataset, random leaves could then be selected to generate data for training the morphological classification model.

The features selected for morphology prediction were: color, between light-green, dark-green, yellow-green, and yellow; shape, between ovate, lanceolate, elliptical, and oblong \citep{nakano2020plant}; and level of brown splotches (representing the level of the withering of the leaf), between none, low, medium, and high.

\begin{center}
    \includegraphics[width=0.45\textwidth]{images/leaf_shapes.jpeg}
    \captionof{figure}{Leaf shape reference from the article by \cite{nakano2020plant}}
\end{center}

By using a slightly modified version of the aforementioned feature extraction function in conjunction with 
a Scikit-Learn \verb|LabelEncoder| to prepare the target features, the classifier could be implemented with Scikit-Learn's \verb|MultiOutputClassifier|. This is a framework for using the same algorithm on three different features at once; namely, the algorithm used was \verb|XGBoost|, which seemed to trump other models in terms of accuracy.

After the model was trained, it was used on each image in the dataset by finding all eligible leaves and determining the mode classification after each eligible leaf was run through the aforementioned model. The results were saved in a copy of the spreadsheet from the previous step.

@@

The classifications were easily attainable because they were saved in a spreadsheet. However, the challenge for using classification data as features to train a model with is that they are categorical, rather than quantitative; data needed to undergo extra transformations to prepare it for use in predictions.

To properly represent the classifications, the method of one-hot encoding \citep{brownlee2020why} was implemented using the Pandas \verb|get_dummies| method. This is a process in which a feature is split into multiple columns, with each row having a one if it matches the column or a zero if it doesn't.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c||c c c c|} 
 \hline
 leaf\_color & light\_green & dark\_green & yellow\_green & yellow  \\
 \hline
 light\_green & 1 & 0 & 0 & 0 \\
 yellow\_green & 0 & 0 & 1 & 0 \\
 dark\_green & 0 & 1 & 0 & 0 \\ 
 light\_green & 1 & 0 & 0 & 0 \\
 yellow & 0 & 0 & 0 & 1 \\
 \hline
\end{tabular}
}
\captionof{table}{Example transformation of a column with one-hot encoding, with the original column on the left and the encoded columns on the right}
\label{Table1}
\end{center}

This process was done with the leaf color and shape variables. However, since a leaf's level of brown splotches is ordinal, it was assigned levels 0, 1, 2, and 3, corresponding to 'none', 'low', 'medium,' and ‘high.' The data for expected treatments were obtained by reading rows of the spreadsheet that had treatments successfully read.

Now, using this data, a predictor could be trained and deployed. The chosen classifier for this task was Scikit-Learn's \verb|RandomForestClassifier|, as it has shown to be a capable model throughout the project. After the model was trained, it was deployed on the spreadsheet, predicting treatments for rows that lacked a properly read one.

@@

First, we aimed to measure the size of a leaf through metadata encoded in EXIF Tags, which are embedded within each of the images. These tags typically give important information about the images, including the type of camera utilized, the distance of the camera from a photo, and photo dimensions (in pixels). In order to initiate the extraction of such metadata from the images, the \verb|exif| package was used and gave a variety of retrieval options.

Out of all the metadata extracted from the EXIF tags, the most notable pieces of information included \verb|ResolutionX|/\verb|ResolutionY| tags and \verb|FocalLength|. Although these pieces of information can assist in developing a mechanism to measure leaf size in the dataset, it simply does not suffice alone; additional information, such as how far the leaf was from the camera, was needed. With more information, such as \verb|FocalPlaneResolution| - showing the number of pixels per real unit (\verb|in| or \verb|cm|) - in addition to previously mentioned features, it is possible to assess the size of specific components of an image. For example, a study on maize phenotyping by \cite{liu2021pocketmaize} showed that embedded information in images was enough for a smartphone application to accurately determine measurements of plant height and leaf area in a photo.

Another approach we considered to measure leaf size included using the white labels, held up by the researcher in each image, as a known reference object. Although this method seemed promising, there were certain limitations in most images in the dataset that prevented a consistent measurement. First, the white label in the image was often obstructed by the researcher’s hands, as shown in \ref{app_blocked}. This obstruction would ultimately interfere with the alignment of the white label, making the label unreliable as an object of reference. Another limitation to this method came from the fact that the labels were often tilted, not in the same plane as the leaves. The tilt of the labels caused inaccuracies in measurement and would lead to misrepresentation of the true length. Along with that, there was no way of finding out how far forward or behind the leaf was relative to the tag, meaning the method would result in an inaccurate representation of the leaf. Again, this issue could have been resolved with more careful documentation of images without obstruction and a better-positioned label alignment.

In the second part of challenge 4, we aimed to find a correlation between phenotypes and environmental features, such as soil type and weather conditions. Using the EXIF tags, we were able to retrieve the coordinates where the image was taken. We noticed all of the images contained were taken in approximately the same two locations, so any weather or soil API would be unable to provide specific conditions. If we were given this information as part of the dataset, we would have been better able to find correlations. Regardless, a study on how the temperature affects the phenotype and development plasticity in Arabidopsis Thaliana \citep{ibanez2017ambient} placed the plants in different temperature environments in order to measure their differences. If we had more information regarding the conditions in which the plants were raised, we would be better able to determine if there was a correlation between phenotypes and environment.

\section{Results and Discussion}
\label{results}

All confusion matrices used in this section originate from Scikit-Learn's \verb|confusion_matrix| and \verb|ConfusionMatrixDisplay|.

@@

The OCR was able to read 91.3\% of the labels in the images of the entire dataset. On the 30 test images, the model had a 77.33\% accuracy on extracted features (94.31\% with null values omitted). A subjective analysis showed that for all features except for genotype, the interpreted values seem to match the true values almost always.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c} 
 \hline
 filename & treatment & block & row & position & genotype \\
 \hline
    ... & D & 1 & 8  & 32 & BESC-34 \\
    ... & C & 1 & 10 & 12 & **BESC-417\_LM**,core \\
    ... & C & 2 & 3  & 40 & BESC-468 \\
    ... & C & 2 & 6  & 54 & BESC-28\_LM \\
    ... & C & 1 & 24 & 22 & **LILD-26-5\_LM**,core \\
 \hline
\end{tabular}
}
\captionof{table}{First five rows of data saved to the spreadsheet from step 1, with filenames omitted to conserve space}
\label{Table1}
\end{center}

Before the text was inserted into the above spreadsheet, it was present in a Pandas \verb|DataFrame|. Here, it was possible to analyze the statistics of the data using the \verb|info| method, containing useful information such as null value count:

\begin{verbatim}
    RangeIndex: 1672 entries, 0 to 1671
    Data columns (total 6 columns):
     #   Column     Non-Null Count  Dtype  
    ---  ------     --------------  -----  
     0   filename   1672 non-null   object 
     1   treatment  1098 non-null   object 
     2   block      1306 non-null   float64
     3   row        1388 non-null   float64
     4   position   1414 non-null   float64
     5   genotype   1431 non-null   object 
\end{verbatim}

Evidently, the former attributes were read the least compared to those near the end of the tag, with 66\% non-null values for the treatment and 86\% non-null values for the genotype. This is likely due to the many forms of obstructions in the images, such as a leaf, a data collector's hand, or even the boundary of the image itself. Examples of such can be seen in \ref{app_blocked}.

Even though the information suggests that there is more error in reading the treatment, it is more likely that it stems from text processing. The interpreted genotype has more room for error, as there is an extensive number of possibilities for it, while the treatment must be a \verb|C| or a \verb|D|, leaving little room for error.

Regardless, it is undeniable that the use of \verb|PaddleOCR|, a relatively new and underused model, proved significantly beneficial and successful in reading labels from the images in the dataset. However, one evident issue lies in the fact that the suffix of the genotype (such as \verb|_LM|) was often not read properly, causing the text processing to drop it entirely. Increased text pre-processing may have alleviated this issue.

@@

The spreadsheet from the last step was successfully updated with morphological predictions from this step.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{c|c|c|c|c|} 
 \hline
 genotype & leaf\_color & leaf\_shape & brown\_splotches \\
 \hline
    BESC-34	&                  light\_green &   ovate &	    none     \\
    **BESC-417\_LM**,core &	   yellow\_green	&  ovate &      high     \\
    BESC-468 &	               yellow &        lanceolate &	high     \\
    BESC-28\_LM &	               dark\_green &	   ovate &      low      \\
    **LILD-26-5\_LM**,core &	   yellow\_green	&  lanceolate &	medium   \\
 \hline
\end{tabular}
}
\captionof{table}{A continuation of Table 2, with morphological data added from step 2}
\label{Table1}
\end{center}

The model had an average accuracy of 62.82\%, which, when considering the fact that each feature had four different classes, is a reasonable accuracy level. The confusion matrices generated by comparing test and predicted data for each feature are below:

\begin{center}
    \includegraphics[width=0.45\textwidth]{images/morph_conf_matrix/color.png}
    \captionof{figure}{Confusion matrix for testing data on the leaf color predictor}
\end{center}

This part of the model had an accuracy score of 69.23\%. It seems to be reliable in predicting leaf color but has difficulty classifying true yellow-green leaves; however, it still correctly predicts yellow-green most of the time. Because the testing data seems to lack yellow leaves, the model's accuracy regarding them is unknown.

\begin{center}
    \includegraphics[width=0.45\textwidth]{images/morph_conf_matrix/shape.png}
    \captionof{figure}{Confusion matrix for testing data on the leaf shape predictor}
\end{center}
